---
title: "L&T credit analysis"
output: html_document
date: "2023-03-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(lubridate)
df <- read_csv("D:/Credit risk/train.csv", col_types = cols(Date.of.Birth = col_date(format = "%d-%m-%y"), DisbursalDate = col_date(format = "%d-%m-%y")))

```
# 1. Cleaning data:
```{r}
colSums(is.na(df))
dim(df)
head(df)
```
Examining the time of the report
```{r}
min(df$DisbursalDate)
max(df$DisbursalDate)
```
```{r}
max(df$Date.of.Birth)
```



Based on disbursalDate column we can the time range for the data is from 2018-08-01 to 2018-10-31, we can calculate age of the borrowers as follow:
```{r}
df$age <- 2019 - year(df$Date.of.Birth)
min(df$age)
```

We can see, minimum age is negative, we need to check for anomolies in date of birth columns
```{r}
filter(df, Date.of.Birth > as.Date("2001-01-01"))
```
Some observation have date of birth in the future, for now treat these observation as Na
```{r}
hist(df$age)
hist(df.rm$age, main = paste("Histogram of" , "customer's age (removed negative values)"), xlab = "age")
hist(abs(df$age), main = paste("Histogram of" , "customer's age (taking absolute values)"), xlab = "age")
```

```{r}

df$age[df$age < 0] <- NA
```

```{r}
df$AVERAGE.ACCT.AGE <- as.numeric(str_extract(df$AVERAGE.ACCT.AGE, "\\d+")) * 12 + as.numeric(str_extract(df$AVERAGE.ACCT.AGE, "\\d+(?=mon)"))
df$CREDIT.HISTORY.LENGTH <- as.numeric(str_extract(df$CREDIT.HISTORY.LENGTH, "\\d+")) * 12 + as.numeric(str_extract(df$CREDIT.HISTORY.LENGTH, "\\d+(?=mon)"))
```

```{r}
filter(df, AVERAGE.ACCT.AGE > 700)
```

```{r}
df$PERFORM_CNS.SCORE.DESCRIPTION <- substr(df$PERFORM_CNS.SCORE.DESCRIPTION, 1, 1)
performance_freq <- data.frame(table(df$PERFORM_CNS.SCORE.DESCRIPTION))
```

```{r}
df$Employment.Type <- replace_na(df$Employment.Type, "Others")
df.rm <- na.omit(df)
dim(df.rm)
```

# 2. Feature egineering

One hot encoding for employment type
```{r}
df.rm$emp_salaried <- ifelse(df.rm$Employment.Type == "Salaried", 1, 0)
df.rm$emp_selfemployed <- ifelse(df.rm$Employment.Type == "Self employed", 1, 0)
```
remove irrelevant columns
```{r}
to_drop <- c("UniqueID", "branch_id", "supplier_id",  "Employee_code_ID", "State_ID", "Current_pincode_ID", "Employment.Type", "Date.of.Birth")

df.rm <- df.rm[, !(colnames(df.rm) %in% to_drop)]
```

# 3. Exploratory data analysis:
# 3.1 Checking target's distribution
```{r}
df.rm %>% group_by(loan_default) %>% summarize(count = n()) %>% ggplot(aes(x = loan_default, y = count)) + geom_col(color = c("red", "green"))
```


# 3.2 
```{r}
write_excel_csv(df.rm, "df_credit_clean2.csv")
dim(df.rm)
```







